A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation

### 2016-08-25 twostagedo.tex
edited abstract, title and introduction.

% The paper headers
\markboth{IEEE CDS}%
{Arriola-Rios \MakeLowercase{\textit{et al.}}: A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
In this paper we present a new framework for learning generative models of object deformation under robotic action. The model is multi-modal in that it is based on integrating force and visual information. Our framework goes beyond previous techniques by simultaneously enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.  While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered them all in a single learning model framework, and none has offered classification from a generative predictable model.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
Deformable objects, prediction, classification, learning.
\end{IEEEkeywords}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A} major challenge in robotic manipulation is to plan actions with objects so as to deform them into new shapes. Most objects that are manipulated by animals in natural environments are deformable, and these include objects that are elastic, plastic, breakable, tearable, and spreadable. This work is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required. These sensorimotor contingencies can then be used to feedback on themselves to predict over long timescales. Without such models planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.

There is a wide range of models of deformation used in engineering, that are capable of producing predictions, such as finite element models, but these typically require careful optimisation by hand over many weeks or months. In graphics there are many models used for generating qualitatively plausible simulations of objects, but these are not necessarily easy to tune to real data. By contrast robots must perform system identification of their model parameters from data, on-line, in reasonable amounts of training time. This includes building models from the usually very limited sensing typically available, such as partial visual views of an object, and force-torque monitoring at contact points. Thus tackling this problem in robotics adds additional difficulties to an already challenging problem.

@@

Previous work on learning sensorimotor contingencies for deformable objects in robotics has been suprisingly limited. Table~\ref{t:advances} summarises the features of the major efforts to date. In each case critical abilities are missing: either models cannot be learned from data; or only address one type of deformation (e.g. elastic); or cannot perform classification; or cannot predict the forces felt; or cannot predict what will happen when the robot loses contact with the object (will it retain or recover from the deformation?). In this paper we present a framework for solving all these problems simultaneously. The key novel contribution we make is to bring together a model of what forces the robot will feel at a contact point with a model of how an object will deform under such applied forces over time. This formulation allows the robot to make predictions about the type of object, and thus it's future behaviour using the resistive forces and the object deformations \fref{fig:forceElasticPlastic} and \fref{fig:sstrain}. A second critical contribution is that our model is both generative and generalisable with respect to contact point.  Finally the third major feature of the model is that it is adaptive: a set of model parameters is learned for each material presented from data, thus creating a model library of materials. This generative property enables both prediction suitable for use in planning, and also classification of materials. The overall model structure is summarised in \fref{fig:system}. Essentially the approach is to use visual tracking of shape deformations to create feedback that is used to train the parameters of a model (Model A) that predicts deformation when coupled with contact forces. The contact forces are themselves predicted using a second model (Model B). In the specific implementation of our system we use a mass spring system for Model A, and a regression model for Model B. These models are learned from the same data, and are both indexable by a known material type from the library. It is important to note, however, that the model structure is separate from the specific model forms used, and that other model types could be substituted for one or both stages.  %This behaviour is described in \alref{alg:overall}.

@@

In summary the two stage model we present gives us the ability to solve {\em simultaneously} for the first time the following problems. Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed. In the rest of the paper we present \comment{an overview of the model in more detail (Section~\ref{sec:model_overview});} related work (Section~\ref{sec:background}); describe the model (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); and the experiments on prediction and classification with both elastic and plastic objects (Section~\ref{sec:experiments}); finishing with a short discussion (Section~\ref{sec:discussion}).

@@

\comment{
A mass-spring model is a very well established abstraction used frequently to simulate the behaviour of some deformable objects.  As a physics based model, once its parameters are properly assigned, its behaviour is determined by the evolution of a set of differential equations.  Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the interacting forces are know.  The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model for the particular object.  For this reason, there are variations that address specific issues, like \cite{Bourguignon2000} and \cite{Teschner2004}.  In this work we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration to a system that exploits it to solve prediction and classification tasks.  We emphasize that the physics based model could be easily substituted by another equivalent model.  For example, the same machine will work if another type of mass-spring model is used, or a FEM model, even though numerical results will differ to some extent.  For these experiments we chose to use a slightly modified version of the mass-spring model proposed by Teschner in \cite{Teschner2004}.

@@

In a simple mass-spring model, the shape of an object is approximated by a uniform geometric mesh, usually made of triangles (2D) or tetrahedrons (3D).  At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located, and is known as a \textit{mass particle}.  The edges between vertices roughly model the interactions between particles, that attract each other up to a certain critical distance, but repel each other if they become closer, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material over itself.  The simplest model associates a linear spring with each edge, as if those forces would occur only between connected neighbours.  The resting state of the solid corresponds to a configuration of the masses and springs where the sum of all forces acting over each particle is zero, and they are said to be in equilibrium.

@@

The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004}, which is explained below, was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.  The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.  Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.  The potential energy terms for the preservation of length with damping and preservation of area were included in our system \comment{exactly as they were defined by Teschner}.  To compensate for the lack of a preservation of volume, which tends to avoid distortions of the graph \comment{in the original model}, \comment{we derived } a new term for the preservation of the angles \comment{in every triangle} \sout{was included}.  The relevant equations are presented in the following paragraphs, while detailed derivations can be found in \cite{Arriola2013thesis}.

\comment{Teschner defined a generalized version of the simple mass-spring model by using the concept of \emph{constraint}.  He makes use of different constraints to model the tendency of an object to recover several of its attributes.} For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.  When the value of the constraint is zero, the potential energy is zero as well.  When the constraint is not satisfied, the potential energy increases.  This behaviour is obtained through the following definition for the energy:
\begin{align}
 E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint}
\end{align}
where $k$ is a proportionality constant \comment{\sout{\eref{eq:EConstraint}}} and $p_i$ is the position of particle $i$.

In order to reduce the energy of the system when a constraint is not satisfied, the force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
\begin{align}
 \vec{F}^i(\vec{p}_1,...,\vec{p}_{n}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
\end{align}
Given the oscilatory nature of these equations, \comment{when using constraints that model the behaviour of springs,} damping can be introduced to stop the oscillations.  This is modelled with a force acting in the opposite direction to the velocity of the affected particles\comment{, thus slowing down motion.  Teschner wrote this as a function of the constraints as well \cite{Teschner2004}}:
\begin{align}
 \label{eq:Fdamping}
 \vec{F}^i&(\vec{p}_1,...,\vec{p}_{n},\vec{v}_1,...,\vec{v}_{n}) = \nonumber \\
    &\left( -kC-k_d\sum_{1 \le j \le n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
 \dfrac{\partial C}{\partial \vec{p}_i},
\end{align}
where $\vec{v}_j = \frac{\partial \vec{p}_j}{\partial t}$ is the velocity of the $j^{th}$ particle connected to $i$, when the constraint acts on $n$ particles.

\subsubsection{Preservation of Length}
The constraint for the potential energy is given by:
\begin{equation}
 C = \dfrac{|\vec{p}_j-\vec{p}_i|-D_0}{D_0}, \label{eq:clength}
\end{equation}
where $\vec{p}_i$ and $\vec{p}_j$ are connected by a spring and $D_0$ is the rest length of that spring.  The force derived from this constraint pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].  Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent \cite{Teschner2004}.  Morris \cite{Morris2008} omitted that in his final equations for the force.  Here we use the normalised force $\vec{F}^i_D$ \eref{eq:FDlength}, which has the effect that, if we obtain a good set of parameters for a certain mesh of high resolution, the same set can be used for a mesh of lesser resolution that preserves the same type of symmetries.  Therefore, substituting \eref{eq:clength} in \eref{eq:Fdamping}, after some manipulation we get:
\begin{align}
 \vec{F}^i_D(p_i,p_j,v_i,v_j) &= \dfrac{k_L}{D_0^2} \left( |\vec{p}_j-\vec{p}_i|-D_0 \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|}  + \nonumber \\
 & \dfrac{k_d}{D_0^2}\left( \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \cdot (\vec{v}_j-\vec{v}_i) \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \label{eq:FDlength}
\end{align}
where $k_L$ is the linear spring constant and $k_d$ the damping constant.

\subsubsection{Preservation of Area}
This constraint is applied per triangle in the mesh.  Since Teschner and Morris \comment{\sout{didn't find it helpful to add damping to the preservation of areas, it is avoided here.} found that the use of damping for the preservation of areas does not improve significantly the stability of the simulation, it is not included here either.}  Therefore the force $\vec{F}^i_A$ is derived as follows:
\begin{align}
 C &= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0} \\
 Area &= \frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|, \\
 \vec{F}^i_A(p_i,p_j,p_k) &= -\frac{k_A}{2A_0^2} \left( Area-A_0 \right) \nonumber \\
 & \frac{(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)}{|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|} [\vec{1} \times (\vec{p}_j-\vec{p}_k)] \label{eq:area_force}
\end{align}
where $A_0$ is the initial area of the triangle and $k_A$ the proportionality constant.

@@

\subsubsection{Preservation of Angles}
The previous terms can not do anything to restore the mesh if triangles get flipped during deformation.  For this reason we included a new term that enforces the preservation of the original angles of the triangles.  [\fref{fig:forces}(c)] gives an idea of how these forces look.  Here the energy depends on the difference \comment{\sout{of the angles between adjacent edges.} between the current angle, between adjacent edges, and the angle between them at the equilibrium position.}

@@

\subsection{Plastic Deformation}
\begin{figure}
 \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
 %\STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
 \comment{\STATE Let $l$ be the length of the edge at time $t+h$, after an integration step where new positions of its vertices $p_i(t+h)$ where calculated.} And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
 \IF {$\alpha > max\_\alpha$}
  \STATE $l_0 \gets l$
 \ELSE
  \STATE $l_0 \gets (1.0 - \alpha) D_0$
 \ENDIF
 \STATE $elastic\_deformation \gets \frac{l_0 - l}{D_0}$
 \IF {$elastic\_deformation > yield$}
  \STATE $\alpha \gets \alpha + creep * elastic\_deformation$
  \IF {$\alpha > max\_\alpha$}
   \STATE $\alpha \gets max\_\alpha$
  \ENDIF
 \ENDIF
 \end{algorithmicieee}
\caption{Algorithm to calculate the permanent deformation of an edge.}
\end{figure}

@@

\subsection{Integration Scheme}
\label{sec:ischeme}
Rather than using the traditional Newton equations with a numeric integration scheme like Euler \comment{or} Verlet, \comment{used twice to solve the second order differential equiations,} we made the \comment{\sout{velocities proportional to the forces:} force proportional to the velocity:
\begin{align}
 F(t) = m \dfrac{\partial \vec{p}}{\partial t}
\end{align}
and applied Euler only once to solve for $p$:
\begin{equation}
 \vec{p}(t+h) = \vec{p}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
\end{equation}}
with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{p}(t+h)$, the position of the vertex at time $t+h$.  \comment{We acknowledge that this is not the proper Newtonian model, however t}he simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
When experimenting with different integration steps, force values were interpolated between measurements.

>>>>

% The paper headers
\markboth{IEEE CDS}%
{Arriola-Rios \MakeLowercase{\textit{et al.}}: Multi-Modal Modular Sensorimotor Contingencies for Deformable Objects}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
\comment{In this paper we present a modular, multi-modal framework for learning generative models of object deformation under robotic pushing. The model is multi-modal in that it is based on integrating force and visual information. The framework is also modular, being composed of a learned module for each type of material. Each such module is learned by parameter estimation for a general model form. When presented with a test example--a robot finger pushing a deformable object made of an unidentified, but previously learned, material--the predictions of the modules are compared so as to classify the material.} Our framework goes beyond previous techniques by simultaneously enabling i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from both force and visual data, v) prediction of object behaviour after contact by the robot terminates.  While previous work on deformable object behaviour in robotics has offered one or two of these features none has offered them all in a single learning framework, and none has offered classification from a generative model.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
deformable objects, prediction, classification, learning.
\end{IEEEkeywords}

@@

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{A} major challenge in robotic manipulation is to plan actions with objects so as to deform them into new shapes. Most objects that are manipulated by animals in natural environments are deformable, including objects that are elastic, plastic, breakable, tearable, and spreadable. \comment{The work reported here is based on the premise that in order to plan manipulations of deformable objects, the sensorimotor contingencies governing object deformation are required. A sensorimotor contingency is simply defined as the relationship between the robot's actions and the resultant changes in its observations. Thus different materials (elastic, plastic) present have different sensorimotor contingencies. The main hypothesis is that predictive models of these sensorimotor contingencies can be learned from data. These learned models can then be used to make the predictions required for planning. The predictions of a learned model can be fed back into the model, to predict an object's deformation behaviour over a variety of time scales. Without such predictions planners cannot reason about the effects of actions, or sequences of actions, to achieve desired effects.

Various models of deformation are used in engineering, such as finite element models, but these typically require careful optimisation by hand, and cannot run in real time. In graphics there are many methods used for generating qualitatively plausible simulations of deformable objects, but these are not easy to fit to real data. In contrast, robots must learn model parameters from data, on-line, in reasonable amounts of training time. This includes learning from the limited sensing typically available, such as partial visual views of an object, and force-torque monitoring only at contact points. Thus, tackling the prediction problem in robotics adds additional difficulties to an already challenging problem.}

@@

\comment{Previous work on learning sensorimotor contingencies for deformable objects in robotics has been limited. Table~\ref{t:advances} summarises the features of the major efforts to date. In each case some abilities are missing: either models cannot be learned from data; or only address one type of deformation (e.g. elastic); or cannot perform classification; or cannot predict the forces felt; or cannot predict what will happen when the robot loses contact with the object (will it retain or recover from the deformation?). In this paper we present a framework for solving all these problems simultaneously.

There are several novel contributions. First, we show how the modular learning framework of \cite{harunowolpert,kopickiwyatt16} can be applied to deformable objects. Second, we show how each module can be decomposed into two separate learning problems, one for predicting forces from finger motions, and one for predicting object deformations from forces. Both models are generative, and we show how to learn each from data. Third, we show how to use these learned generative models not only to predict deformation, but also to classify an unidentified material.}

 %This behaviour is described in \alref{alg:overall}.

@@

In summary the framework we present gives us the ability to solve {\em simultaneously} for the first time the following problems. Our model enables i) predictions over many steps, ii) learning of plastic and elastic deformation from real data, iii) prediction of forces experienced by the robot, iv) classification of materials from either or both force and visual data, v) prediction of object behaviour after contact by the robot is removed. In the rest of the paper we present \comment{an overview of the model in more detail (Section~\ref{sec:model_overview});} related work (Section~\ref{sec:background}); describe the model (Section~\ref{sec:model}); describe how the model parameters are learned from real data (Section~\ref{sec:learning}); and the experiments on prediction and classification with both elastic and plastic objects (Section~\ref{sec:experiments}); finishing with a short discussion (Section~\ref{sec:discussion}).

@@

\comment{
A mass-spring model is a very well established abstraction used to emulate the behaviour of deformable objects.  As a physics based model, once its parameters are properly assigned, its behaviour is determined by the evolution of a set of differential equations.  Therefore, by offering a method to automatically calibrate those parameters, this model can make long term predictions about the behaviour of deformable objects, under previously unseen interactions, if only the interacting forces are know.  The range of validity of these predictions greatly depends on the quality of the calibration, but also on the suitability of the mass-spring model for the particular object.  For this reason, there are variations that address specific issues, like \cite{Bourguignon2000} and \cite{Teschner2004}.  In this work we focus on the machinery required to achieve this automatic calibration of a physics based model and its integration to a system that exploits it to solve prediction and classification tasks.  We emphasize that the physics based model could be easily substituted by another equivalent model.  For example, the same machine will work if another type of mass-spring model is used, or a FEM model, even though numerical results will differ to some extent.  For these experiments we chose to use a slightly modified version of the mass-spring model proposed by Teschner in \cite{Teschner2004}.

@@

In a simple mass-spring model, the shape of an object is approximated by a uniform geometric mesh, usually made of triangles (2D) or tetrahedrons (3D).  At every vertex $i$ in the mesh, an ideal particle with no volume, but with mass $m_i$, is located.  The edges between vertices roughly model the interactions between particles that attract each other and at the same time repel each other, therefore maintaining the cohesion of the solid, while avoiding the collapse of the material over itself.  The simplest model associates a linear spring with each edge, as if those forces would occur only between connected neighbours.  The resting state of the solid corresponds to a configuration of the masses and particles where the sum of all forces acting over each particle is zero and they are said to be in equilibrium.

@@

The mass-spring model proposed by Teschner et al.\ in \cite{Teschner2004} was reduced to a 2D model, so that the predictions could be compared with 2D images of the deformable object.  The main idea being that an agent can make rough real time predictions of what it can see and touch and use those predictions to solve various tasks, like classification or grasping.  Here, the surface of the deformable object is approximated by a regular triangular mesh of unitary masses and springs.  The potential energy terms for the preservation of length with damping and preservation of area were included in our system \comment{exactly as they were defined by Teschner}.  To compensate for the lack of a preservation of volume, which tends to avoid distortions of the graph \comment{in the original model}, \comment{we derived } a new term for the preservation of the angles \comment{in every triangle} \sout{was included}.  The relevant equations are presented in the following paragraphs, while detailed derivations can be found in \cite{Arriola2013thesis}.

\comment{Teschner defined a generalized version of the simple mass-spring model by using the concept of \emph{constraint}.  He makes use of different constraints to model the tendency of an object to recover several of its attributes.} For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.  When the value of the constraint is zero, the potential energy is zero as well.  When the constraint is not satisfied, the potential energy increases.  This behaviour is obtained through the following definition for the energy:
\begin{align}
 E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint}
\end{align}
where $k$ is a proportionality constant \sout{\eref{eq:EConstraint}} and $p_i$ is the position of particle $i$.

In order to reduce the energy of the system when a constraint is not satisfied, the force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
\begin{align}
 \vec{F}^i(\vec{p}_1,...,\vec{p}_{n}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
\end{align}
Given the oscilatory nature of these equations, damping can be introduced to stop the oscillations.  This is modelled with a force acting in the opposite direction to the velocity of the affected particles:
\begin{align}
 \label{eq:Fdamping}
 \vec{F}^i&(\vec{p}_1,...,\vec{p}_{n},\vec{v}_1,...,\vec{v}_{n}) = \nonumber \\
    &\left( -kC-k_d\sum_{1 \le j \le n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
 \dfrac{\partial C}{\partial \vec{p}_i},
\end{align}
where $\vec{v}_j = \frac{\partial \vec{p}_j}{\partial t}$ is the velocity of the $j^{th}$ particle connected to $i$, when the constraint acts on $n$ particles.

\subsubsection{Preservation of Length}
The constraint for the potential energy is given by:
\begin{equation}
 C = \dfrac{|\vec{p}_j-\vec{p}_i|-D_0}{D_0}, \label{eq:clength}
\end{equation}
where $\vec{p}_i$ and $\vec{p}_j$ are connected by a spring and $D_0$ is the rest length of that spring.  The force derived from this constraint pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].  Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent.  Morris \cite{Morris2008} omitted that in his final equations for the force.  Here we use the normalised force $\vec{F}^i_D$ \eref{eq:FDlength}, which has the effect that, if we obtain a good set of parameters for a certain mesh of high resolution, the same set can be used for a mesh of lesser resolution that preserves the same type of symmetries.  Therefore, substituting \eref{eq:clength} in \eref{eq:Fdamping}, after some manipulation we get:
\begin{align}
 \vec{F}^i_D(p_i,p_j,v_i,v_j) &= \dfrac{k_L}{D_0^2} \left( |\vec{p}_j-\vec{p}_i|-D_0 \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|}  + \nonumber \\
 & \dfrac{k_d}{D_0^2}\left( \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \cdot (\vec{v}_j-\vec{v}_i) \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \label{eq:FDlength}
\end{align}
where $k_L$ is the linear spring constant and $k_d$ the damping constant.

\subsubsection{Preservation of Area}
This constraint is applied per triangle in the mesh.  Since Teschner and Morris didn't find it helpful to add damping to the preservation of areas, it is avoided here.  Therefore the force $\vec{F}^i_A$ is derived as follows:
\begin{align}
 C &= \frac{\frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|-A_0}{A_0} \\
 Area &= \frac{1}{2}|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|, \\
 \vec{F}^i_A(p_i,p_j,p_k) &= -\frac{k_A}{2A_0^2} \left( Area-A_0 \right) \nonumber \\
 & \frac{(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)}{|(\vec{p}_j-\vec{p}_i)\times(\vec{p}_k-\vec{p}_i)|} [\vec{1} \times (\vec{p}_j-\vec{p}_k)] \label{eq:area_force}
\end{align}
where $A_0$ is the initial area of the triangle and $k_A$ the proportionality constant.

@@

\subsubsection{Preservation of Angles}
The previous terms can not do anything to restore the mesh if triangles get flipped during deformation.  For this reason we included a new term that enforces the preservation of the original angles of the triangles.  [\fref{fig:forces}(c)] gives an idea of how these forces look.  Here the energy depends on the difference of the angles between adjacent edges.

@@

\subsection{Plastic Deformation}
\begin{figure}
 \begin{algorithmicieee}{Plastic Deformation of an Edge}\label{alg:deformEdge}
 \STATE Let $l$ be the new length of the edge, after an integration step. And $max\_\alpha$ the maximum proportional permanent deformation being allowed.
 \IF {$\alpha > max\_\alpha$}
  \STATE $l_0 \gets l$
 \ELSE
  \STATE $l_0 \gets (1.0 - \alpha) D_0$
 \ENDIF
 \STATE $elastic\_deformation \gets \frac{l_0 - l}{D_0}$
 \IF {$elastic\_deformation > yield$}
  \STATE $\alpha \gets \alpha + creep * elastic\_deformation$
  \IF {$\alpha > max\_\alpha$}
   \STATE $\alpha \gets max\_\alpha$
  \ENDIF
 \ENDIF
 \end{algorithmicieee}
\caption{Algorithm to calculate the permanent deformation of an edge.}
\end{figure}

@@

\subsection{Integration Scheme}
\label{sec:ischeme}
Rather than using the traditional Newton equations with a numeric integration scheme like Euler of Verlet, we made the velocities proportional to the forces:
\begin{equation}
 \vec{x}(t+h) = \vec{x}(t) + h \dfrac{\vec{F(t)}}{m}\label{eq:integration}
\end{equation}
with $\vec{F}(t) = \vec{F}_D(t) + \vec{F}_A(t) + F_\varphi(t)$ being the sum of the forces for the preservation of length, area and angle respectively and $\vec{x}(t+h)$, the position of the vertex at time $t+h$.  The simulations obtained were closer to the observed behaviour, because there is not inertia in the oscillation.
When experimenting with different integration steps, force values were interpolated between measuremente.
